CYO Project Report - Delivery Prediction System

### 2021-07-21 finalProjectReport.Rmd
fine tuned files

---
title: "CYO Project Report - Delivery Prediction System"
author: "Asham Vohra"
date: "6/21/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
---

@@

@@

This was a crucial dataset not only for us to build our delivery estimation system but to even identify the current performance metric i.e. RMSE for the approach Olist was using.

@@

```{r Data wrangling-order items data set, include=FALSE, echo=FALSE}
# orders data with item related seller info
# order_item_id --->sequential number identifying number of items included in the same order.
# so if there are three items they will be nuber 1 , 2, 3 for the given order
# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))  %>% select(-order_item_id)

@@

For the above analysis, knn model was used, given that it does not make assumptions about the dataset, which was useful in the early phase of analysis. However, to use knn model, we **scaled and normalized** the features as knn model uses Euclidean distance. Scaling and normalization was carried our using training data and then used for test data.

@@

In addition, for knn model training and tuning, we carried our 5-fold cross validation providing a range of values of k to use for training the model. Here, k means number of neighbors to be considered. The cross validation helped in removing any randomness from our tuning parameters. In addition, this helped reduce the probability of overfitting.

@@

Based on model results, k value of `r if(exists("knn_fit",mode="list")) { knn_fit$finalModel$k } else {77}` was identified to be the one with lowest RMSE.

@@

To train and tune random forest model, we carried our 4-fold cross validation providing a range of values of mtry to use for training the model. Here, mtry means number of random features used to determine the best split at each node of the decision tree. In addition, we configured the number of decision trees to create as 200. Here, the cross validation not only helped to reduce the probability of overfitting but also helped in identifying the tuning parameter i.e. mtry which would work across randomly selected datasets.

@@

Based on model results, mtry value of `r if(exists("rf_fit",mode="list")) { rf_fit$finalModel$mtry } else {2}` was identified to be the one with lowest RMSE.

@@

We noted that weighted ensemble model had lowest RMSE of `r min(rmses_weighted_ensemble_model)` at weight=`r weights_for_ensemble_model[which.min(rmses_weighted_ensemble_model)]`

>>>>

---
title: "CYO Project Report - Delivery Prediction System"
author: "Asham Vohra"
date: "21st June 2021"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
  html_document:
    df_print: paged
urlcolor: blue
---

@@

\newpage

@@

This was a crucial dataset not only for us to build our delivery prediction system but to even identify the current performance metric i.e. RMSE for the approach Olist was using.

@@

```{r Data wrangling-order items data set, include=FALSE, echo=FALSE}
# orders data with item related seller info
# order_item_id --->sequential number identifying number of items included in the same order.
# so if there are three items they will be numbered 1 , 2, 3 for the given order
# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
# freight_value --> item freight value item (if an order has more than one item the freight value is split between items)
order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))  %>% select(-order_item_id)

@@

For the above analysis, k-nearest neighbors(knn) model was used, given that it does not make assumptions about the dataset, which was useful in the early phase of analysis. However, to use knn model, we **scaled and normalized** the features as knn model uses Euclidean distance. Scaling and normalization was carried our using training data and then used for test data.

@@

In addition, for knn model training and tuning, we carried our 5-fold cross validation providing a range of values of **k** to use for training the model. Here, k means number of neighbors to be considered. The cross validation helped in removing any randomness from our tuning parameters. In addition, this helped reduce the probability of overfitting.

@@

Based on model results, **k** value of **`r if(exists("knn_fit",mode="list")) { knn_fit$finalModel$k } else {77}`** was identified to be the one with lowest RMSE.

@@

To train and tune random forest model, we carried our 4-fold cross validation providing a range of values of mtry to use for training the model. Here, **mtry** means number of random features used to determine the best split at each node of the decision tree. In addition, we configured the number of decision trees to create as **200**. Here, the cross validation not only helped to reduce the probability of overfitting but also helped in identifying the tuning parameter i.e. mtry which would work across randomly selected datasets.

@@

Based on model results, **mtry** value of **`r if(exists("rf_fit",mode="list")) { rf_fit$finalModel$mtry } else {2}`** was identified to be the one with lowest RMSE.

@@

We noted that weighted ensemble model had lowest RMSE of **`r min(rmses_weighted_ensemble_model)`** at weight=**`r weights_for_ensemble_model[which.min(rmses_weighted_ensemble_model)]`**
